---
title: "Task 2"
author: "Yvonne Dinh"
date: "12/7/2020"
output: html_document
---

```{r setup, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(tibble)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```

#### Political and Environmental Self-Identification

Introduction: This analysis uses survey data conducted by Chelsea Batavia and the Andrews Forest LTER in August 2017. It explores the association between survey respondents' political affiliation and whether or not they self-identify as an environmentalist. This will be explored through a chi-square test to analyze if the two responses are independent of each other.

```{r, echo=FALSE, warnings=FALSE, include=FALSE}
#read in data
survey=read.csv(here("data","conservation_survey.csv")) %>% 
  clean_names()

#keep only the 'ENV_ID' and 'POL_OR' variables
survey_simple=survey %>% 
  select(env_id, pol_or)

#add a new column of political parties and get rid of other political affiliations
survey_simple=survey_simple %>% 
  mutate(party=case_when(
    pol_or %in% c("1", "2", "3") ~ "Conservative",
    pol_or %in% c("5", "6", "7") ~ "Liberal")) %>% 
  mutate(envi=case_when(
    env_id %in% c("1") ~ "Yes",
    env_id %in% c("2") ~ "No",
    env_id %in% c("3") ~ "Unsure")) %>% 
  filter(!is.na(party))
```

Table 1: Counts and proportions of liberal and conservative survey respondents to the statement "I consider myself an environmentalist." Data: Andrews et al. (2019)
```{r, warnings = FALSE, echo = FALSE}
#create a table that shows counts and proportions of survey results
survey_counts=survey_simple %>% 
  tabyl(party, envi)

survey_table=survey_simple %>%
  tabyl(party, envi) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits=2) %>% 
  adorn_ns() %>% 
  rename ("Political Party" = party) %>% 
  select(1,4,2,3)

#kbl table
survey_table %>% 
  kbl() %>%
  kable_material(c("hover"))

```

```{r,echo=FALSE, warnings=FALSE, include=FALSE}
#convert counts table to a contingency table and preform a chi-square test for independence to answer: is there an association between political affiliation and environmentalism?

#convert counts table to a contingency table
survey_ct=survey_counts %>% 
  column_to_rownames(var="party")

#preform a chi-square test
survey_chisq=chisq.test(survey_ct)
survey_tidy=tidy(survey_chisq)

#is there an association between political affiliation and environmentalism? the p-squared value is significant, so we reject the null hypothesis that there is not independence between the two values. 

```

A majority of liberal respondents (53.73%, n=281) identified as an environmentalist, while a majority of conservative respondents (50.24%, n=318) did not. After running a $\chi$^2^ test, we find that there is a significant association between survey respondents' political affiliation and whether they identify as environmentalists ($\chi$^2^(`r survey_tidy$parameter`) = `r round(survey_tidy$statistic,2)`, *p* = `r signif(survey_tidy$p.value, 3)`). These results indicate that liberals are more likely to identify as environmentalists than conservatives are. 


Data: 

Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c
