---
title: "Task 3"
author: "Yvonne Dinh"
date: "12/7/2020"
output: html_document
---
### Comparison of Desert Grassland and Colorado Checkered female lizard lengths

Introduction: This report uses data on lizards from the Jornada Basin LTER in Southern New Mexico. It compares the mean length of female Desert Grassland and Colorado Checkered lizards using a two-sample t-test and Cohen's d test. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(tibble)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
library(ggbeeswarm)
library(effsize)
```

```{r,echo=FALSE, warnings=FALSE, include=FALSE}
#read in data
lizards=read.csv(here("data","lizards.csv")) %>% 
  clean_names()

#simplify and filter the data and create a new column with renamed species names
liz_simple=lizards %>% 
  select(spp, sex, total_length) %>% 
  filter(sex=="F") %>% 
  filter(spp %in% c("CNUN", "CNTE")) %>% 
  mutate(species = case_when(
    spp=="CNUN" ~ "Desert Grassland Lizard",
    spp=="CNTE" ~ "Colorado Checkered Lizard")
  )
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#create a finalized visualization, i chose a boxplot
ggplot(liz_simple, aes(x=species, y=total_length))+
  labs(x="Species", y="Lizard Length, mm", title="Lengths of Female Lizards")+
  geom_boxplot(color="gray39")+
  geom_beeswarm(color="deeppink3", alpha=0.6)+
  theme_minimal()
```

Figure 1. Boxplot of lizard length for female Colorado Checkered and Desert Grassland lizards. Data: Lightfoot (2017).


Table 1. Descriptive statistics (mean, standard deviation, and sample size) for length of female Colorado Checkered and Desert Grassland lizards. Data: Lightfoot (2017).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#prepare a finalized summary table w/ mean and stdev for total length, sample size, for each species
liz_table=liz_simple %>%
  group_by(species) %>% 
  summarize(mean_length=mean(total_length, na.rm = TRUE),
            sd_length=sd(total_length, na.rm = TRUE), #standard dev
            n_length=n())
#rename columns
colnames(liz_table)=c("Species", "Mean Length, mm", "Standard Deviation, mm", "Sample Size")
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
#kbl to make it nicer looking
liz_table %>% 
  kbl() %>%
  kable_material(c("hover"))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#exploratory analyses

#histogram
ggplot(liz_simple, aes(x=total_length))+
  geom_histogram()+
  facet_wrap (~species)

#qq-plot
ggplot(liz_simple, aes(sample=total_length))+
  geom_qq()+
  facet_wrap(~species)
```

```{r,echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#calculate cohen's d
cnun_lizards=liz_simple %>% 
  filter(spp == "CNUN") %>% 
  filter(!is.na(total_length)) %>% 
  pull(total_length)

cnte_lizards=liz_simple %>%  
  filter(spp == "CNTE") %>% 
  filter(!is.na(total_length)) %>% 
  pull(total_length)

cohen.d(cnun_lizards, cnte_lizards)

#t-test too
tt_liz=t.test(cnun_lizards, cnte_lizards)
tt_clean=broom::tidy(tt_liz)

#find difference in means
mean_diff=liz_table[1,2] - liz_table[2,2]
```

The actual difference in means between the lengths of female Desert Grassland and Colorado Checkered Lizards is `r round(mean_diff,4)` mm and the effect size using Cohen's d is large at -2.42. After running a two-sample t-test, the p-value is *p* = `r signif(tt_clean$p.value, 2)`, meaning that there is a `r signif(tt_clean$p.value, 4)` probability of getting two sample means with the same mean difference we observed. Since the p-value is really small, it is very unlikely that these samples were drawn from populations with the same mean.

Data:

Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 


